---
title: "Comparing {simaerep} and Funnel Plot Performance"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    collapse: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, message = FALSE)
```

# Load
```{r load}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(furrr))
suppressPackageStartupMessages(library(future))
suppressPackageStartupMessages(library(simaerep))

plan(multisession, workers = 6)
```

# Install {gsm}

```{r eval = FALSE}
devtools::install_github("Gilead-BioStats/gsm@v1.9.2", ref = "main")
```



# Introduction

The [{gsm}](https://github.com/Gilead-BioStats/gsm/) R package provides a standardized Risk Based Quality Monitoring (RBQM) framework for clinical trials that pairs a flexible data pipeline with robust reports. It also uses Funnel Plots to flag outliers which provide broader tolerance limits for sites with low exposure and narrower limits for sites with higher exposure. This method is different to the event rate based limits we have used in previous heuristics to measure {simaerep} performance. Funnel plots are discussed in greater detail by [Zink et al. 2018](https://doi.org/10.1177/2168479017738981)

One of the draw backs of using funnel plots for flagging is that they assume that the AE rate remains constant over the course of the study.

# Prepare Data

## Load Portfolio Configurations

We have prepared a snapshot of the AE reporting configuration of our current portfolio. For each study we have also measured a visit-specific AE rate which allows us to generate a synthetic portfolio with flexible AE rates across a study.

```{r real_config}

df_config <- readr::read_csv("ae_conf_20240220.csv")
df_ae_rates <- readr::read_csv("ae_rates_20240220.csv")

df_config %>%
  head(25) %>%
  knitr::kable()

df_ae_rates %>%
  head(25) %>%
  knitr::kable()

```

## Simulate Portfolio

We generate two synthetic portfolios with no AE under-reporting sites. One portfolio with a fixed AE rate for all visits and another one with a flexible visit-specific AE rate.

```{r sim_portf}

df_portf_fix <- sim_test_data_portfolio(df_config, parallel = TRUE, progress = TRUE)

df_portf_fix %>%
  head(25) %>%
  knitr::kable()


df_portf_flex <- sim_test_data_portfolio(df_config, df_ae_rates = df_ae_rates, parallel = TRUE, progress = TRUE)
```

## Compare AE rates

Next we confirm the different AE rates in our two synthetic portfolios.

```{r}
df_rate_fix <- df_portf_fix %>%
  mutate(ae_rate = coalesce(n_ae - lag(n_ae), n_ae), .by = c("study_id", "patnum")) %>%
  summarise(ae_rate = mean(ae_rate), .by = c("study_id", "visit")) %>%
  mutate(rate = "fix")


df_rate_flex <- df_portf_flex %>%
  mutate(ae_rate = coalesce(n_ae - lag(n_ae), n_ae), .by = c("study_id", "patnum")) %>%
  summarise(ae_rate = mean(ae_rate), .by = c("study_id", "visit")) %>%
  mutate(rate = "flex")
```

```{r fig.width=10, fig.height = 7}
bind_rows(df_rate_flex, df_rate_fix) %>%
  ggplot(aes(visit, ae_rate)) +
    geom_line(aes(group = study_id), alpha = 0.2) +
    geom_smooth() +
    facet_wrap(~ rate) +
    labs(title = "Average AE rates per Study")
```


```{r fig.width=10, fig.height = 7}
bind_rows(df_rate_flex, df_rate_fix) %>%
  filter(dense_rank(study_id) <= 16) %>%
  ggplot(aes(visit, ae_rate)) +
    geom_line(aes(group = rate, color = rate)) +
    facet_wrap(~ study_id, scales = "free") +
    labs(title = "Average AE rates for Selected Studies")

```

We can confirm that the AE rates in the "flexible" portfolio are not constant. Moreover
we see that the AE rate profile is very unique for each study.


# Funnel Plots {gsm}

## Example

Here we demonstrate how to use the {gsm} package on our simulated portfolios so that we can get a good visualisation of the funnel plot.

```{r}

get_SUBJ <- function(df_portf) {
  df_portf %>%
    select(study_id, siteid = site_number, subjid = patnum, timeonstudy = visit) %>%
    summarise(timeonstudy = max(timeonstudy), .by = c(study_id, siteid, subjid)) %>%
    group_by(study_id) %>%
    nest()
}


get_AE <- function(df_portf) {
  df_portf %>%
    select(study_id, subjid = patnum, n_ae) %>%
    summarise(n_ae = max(n_ae), .by = c(study_id, subjid)) %>%
    filter(n_ae > 0) %>%
    mutate(n_ae = map(n_ae, ~ tibble(n = seq(1, .)), .progress = TRUE)) %>%
    unnest(n_ae) %>%
    select(- n) %>%
    group_by(study_id) %>%
    nest()
}

dfSUBJ_fix <- get_SUBJ(df_portf_fix)
dfAE_fix <- get_AE(df_portf_fix)

```

```{r}
dfInput <- gsm::AE_Map_Raw(list(dfSUBJ = dfSUBJ_fix$data[[1]], dfAE = dfAE_fix$data[[1]]))
dfInput

dfTransformed <- gsm::Transform_Rate(
  dfInput,
  strNumeratorCol = "Count",
  strDenominatorCol = "Exposure"
  )
dfTransformed

dfAnalyzed <- gsm::Analyze_NormalApprox(dfTransformed)
dfAnalyzed

dfFlagged <- gsm::Flag_NormalApprox(dfAnalyzed, vThreshold = c(-3, -2, 2, 3))
dfFlagged

dfSummary <- gsm::Summarize(dfFlagged)
dfSummary

dfBounds <- gsm::Analyze_NormalApprox_PredictBounds(dfTransformed, vThreshold = c(-3, -2, 2, 3))
dfBounds

chart <- gsm::Visualize_Scatter(dfFlagged, dfBounds)
chart
```

## Simulate UR


# UR Funnel

We write out own funnel function as adapted from {gsm}

```{r}
funnel_ur <- function(df, site, ur_rate) {
  df %>%
    filter(visit == max(visit), .by = patnum) %>%
    summarise(
      Metric = sum(.data$n_ae) / sum(.data$visit),
      n_ae = sum(n_ae),
      visit = sum(visit),
      .by = "site_number"
    ) %>%
    mutate(
      n_ae = ifelse(site_number == site, n_ae * (1 - ur_rate), n_ae),
      Metric = n_ae / visit
    ) %>%
    mutate(
          vMu = sum(.data$n_ae) / sum(.data$visit),
          z_0 = ifelse(.data$vMu == 0,
            0,
            (.data$Metric - .data$vMu) /
              sqrt(.data$vMu / .data$visit)
          ),
          phi = mean(.data$z_0^2),
          z_i = ifelse(.data$vMu == 0 | .data$phi == 0,
            0,
            (.data$Metric - .data$vMu) /
              sqrt(.data$phi * .data$vMu / .data$visit)
          )
    ) %>%
    filter(site_number == site) %>%
    pull(z_i)
}

sim_ur_funnel <- function(df) {
  df %>%
    group_by(study_id) %>%
    nest() %>%
    ungroup() %>%
    mutate(
      sites = map(data, ~ distinct(., site_number))
    ) %>%
    unnest(sites) %>%
    mutate(ur = list(tibble(ur_rate = c(0, 0.1, 0.25, 0.5, 0.75, 1)))) %>%
    unnest(ur) %>%
    mutate(
      score = pmap_dbl(list(data, site_number, ur_rate), funnel_ur, .progress = TRUE)
    )
}

df_sim_ur_funnel_flex <- sim_ur_funnel(df_portf_flex)
```

```{r}
df_sim_ur_funnel_fix <- sim_ur_funnel(df_portf_fix)
```


# UR {simaerep}

We simulate under-reporting for both portfolios using {simaerep} using `sim_ur_scenarios()`.

```{r}
df_sim_simaerep_fix <- sim_ur_scenarios(
   df_portf_fix,
   extra_ur_sites = 0,
   ur_rate = c(0, 0.1, 0.25, 0.5, 0.75, 1),
   parallel = TRUE,
   poisson = TRUE,
   prob_lower = TRUE,
   progress = TRUE
)
```

```{r}
df_sim_simaerep_flex <- sim_ur_scenarios(
   df_portf_flex,
   extra_ur_sites = 0,
   ur_rate = c(0, 0.1, 0.25, 0.5, 0.75, 1),
   parallel = TRUE,
   poisson = TRUE,
   prob_lower = TRUE,
   progress = TRUE
)
```

# Evaluate

## Combine Results

As the funnel plot score does not use multiplicity correction, we also compare the funnel plot score against the {simaerep} score w/o multiplicity correction.

```{r}

df_sim_simaerep_fix$ae_rate <- "AE rate: fix"
df_sim_simaerep_flex$ae_rate <- "AE rate: flexible"
df_sim_ur_funnel_fix$ae_rate <- "AE rate: fix"
df_sim_ur_funnel_flex$ae_rate <- "AE rate: flexible"



df_sim_fun_thresh2 <- bind_rows(df_sim_ur_funnel_fix, df_sim_ur_funnel_flex) %>%
  mutate(
    type = "funnel",
  ) %>%
  select(type, ae_rate, study_id, site_number, ur_rate, score)


df_sim_simaerep_threshp95 <-  bind_rows(df_sim_simaerep_fix, df_sim_simaerep_flex) %>%
  mutate(
    type = "{simaerep}"
  ) %>%
  select(type, ae_rate, study_id, site_number, ur_rate, score = prob_low_prob_ur)

df_sim_simaerep_threshp95_no_mult <-  bind_rows(df_sim_simaerep_fix, df_sim_simaerep_flex) %>%
  mutate(
    type = "{simaerep} no mult"
  ) %>%
  mutate(
    score = 1 - prob_low
  ) %>%
  select(type, ae_rate, study_id, site_number, ur_rate, score)


df_eval <- bind_rows(
  df_sim_simaerep_threshp95,
  df_sim_simaerep_threshp95_no_mult,
  df_sim_fun_thresh2
)
```

## AUC

We continue by comparing the ROC-AUC.

```{r}

get_roc <- function(df_ur, df_nr) {
  df <- bind_rows(df_ur, df_nr) 
  pROC::roc(df, response = "is_ur", predictor = "score", quiet = TRUE)
}

# use 0 scenario to mix with ur scenario and calculate auc from scores
df_nr <- df_eval %>%
  filter(ur_rate == 0) %>%
  mutate(is_ur = "no") %>%
  select(- ur_rate) %>%
  group_by(type, study_id, ae_rate) %>%
  nest() %>%
  ungroup() %>%
  rename(data_nr = data)
  

df_ur <- df_eval %>%
  filter(ur_rate > 0) %>%
  mutate(is_ur = "yes") %>%
  group_by(type, study_id, ur_rate, ae_rate) %>%
  nest() %>%
  ungroup() %>%
  rename(data_ur = data)


df_auc <- df_ur %>%
  left_join(
    df_nr,
    by = c("type", "study_id", "ae_rate")
  ) %>%
  mutate(
    roc = map2(data_ur, data_nr, get_roc, .progress = TRUE),
    auc = map_dbl(roc, pROC::auc, .progress = TRUE)
  ) %>%
  select(type, study_id, ur_rate, ae_rate, roc, auc)

```

### Table

```{r}
df_auc %>%
  summarise(
    sd_auc = sd(.data$auc),
    auc = mean(.data$auc),
    .by = c(type, ur_rate, ae_rate)
  ) %>%
  knitr::kable(digit = 3)
```

### Plot

```{r}

df_auc %>%
  ggplot(aes(type, auc)) +
    geom_boxplot(aes(fill = type)) +
    facet_grid(ae_rate ~ ur_rate)  +
    scale_fill_brewer(palette = "Dark2") +
    theme(axis.text.x = element_blank())

```

## Metrics

### Thresholds

We set the thresholds for funnel and simaerep w/o multiplicity correction so that we get the same fpr as for simaerep with multiplicity correction and the established threshold of 0.95

```{r}

thresh_default <- 0.95

target_fpr <- df_eval %>%
  filter(ur_rate == 0, type == "{simaerep}") %>%
  summarise(fpr = sum(score >= thresh_default) / n()) %>%
  pull(fpr)

thresh_no_mult <-  df_eval %>%
  filter(ur_rate == 0, type == "{simaerep} no mult") %>%
  pull(score) %>%
  quantile(1 - target_fpr)

thresh_funnel <- df_eval %>%
  filter(ur_rate == 0, type == "funnel") %>%
  pull(score) %>%
  quantile(target_fpr)

target_fpr

thresh_no_mult

thresh_funnel
```

### Aggregate

```{r}
get_prop_test_ci95 <- function(..., ix) {
  
  stopifnot(ix %in% c(1, 2))
  
  tryCatch(
    prop.test(...)$conf.int[ix],
    error = function(cnd) c(NA, NA)[ix]
  )
}

aggr_results <- function(df_eval) {

  df_perf <- df_eval %>%
    mutate(
      is_ur = case_when(
        type == "{simaerep}" ~ score >= thresh_default,
        type == "{simaerep} no mult" ~ score >= thresh_no_mult,
        type == "funnel" ~ score <= thresh_funnel
      )
    ) %>%
    summarise(
      n = n(),
      .by = c(type, ae_rate, ur_rate, is_ur)
    ) %>%
    pivot_wider(
      names_from = is_ur,
      values_from = n,
      names_prefix = "is_ur_",
      values_fill = 0
    ) %>%
    mutate(
      n_sites = is_ur_TRUE + is_ur_FALSE,
      ratio = is_ur_TRUE / n_sites,
      ratio_type = ifelse(ur_rate == 0, "fpr", "tpr"),
      ci95_low = map2_dbl(is_ur_TRUE, n_sites, ~ get_prop_test_ci95(.x, .y, ix = 1)),
      ci95_high = map2_dbl(is_ur_TRUE, n_sites, ~ get_prop_test_ci95(.x, .y, ix = 2))
    )
  }

df_perf <- aggr_results(df_eval)
```

### Table

```{r}
df_perf %>%
  knitr::kable(digits = 4)

```

### Plot

```{r, plot, fig.width=10, fig.height = 12}

plot_perf <- function(df_perf) {

  df_perf %>%
    mutate(ur_rate = paste0("under-reporting rate: ",  ur_rate, " - ", ratio_type),
           ur_rate = ifelse(str_detect(ur_rate, "fpr"), "fpr", ur_rate),
           ae_rate = forcats::fct_rev(factor(ae_rate))) %>%
    group_by(ur_rate) %>%
    ggplot(aes(type, ratio)) +
      geom_errorbar(aes(ymin = ci95_low, ymax = ci95_high, color = type, linetype = ae_rate), linewidth = 1) +
      facet_wrap(~ ur_rate, ncol = 1) +
      coord_flip() +
      theme(legend.position = "bottom") +
      labs(
        x = "",
        y = "CI95 Performance Ratio", 
        title = "{simaerep} vs Funnel-Plot Performance"
      ) +
      scale_color_brewer(palette = "Dark2")
}

plot_perf(df_perf)
```


# Summary

- Funnel plot expects constant event rates over time. Performance decrease when event rates are flexible.
- {simaerep} performance is uneffected by flexible event rates.
- {simaerep} detection rates with comparable false positive rates are greater than funnel plot detection rates when event rates are flexible.

```{r close}
plan(sequential)
```
