---
title: "Comparing {simaerep} and {gsm} Performance"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    collapse: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, message = FALSE)
```

# Load
```{r load}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(furrr))
suppressPackageStartupMessages(library(future))
suppressPackageStartupMessages(library(simaerep))
suppressPackageStartupMessages(library(ggExtra))
suppressPackageStartupMessages(library(gridExtra))

plan(multisession, workers = 6)
```

# Install {gsm}

```{r eval = FALSE}
devtools::install_github("Gilead-BioStats/gsm", ref = "main")
```


# Introduction

The [{gsm}](https://github.com/Gilead-BioStats/gsm/) R package provides a standardized Risk Based Quality Monitoring (RBQM) framework for clinical trials that pairs a flexible data pipeline with robust reports. It also uses Funnel Plots to flag outliers which provide broader tolerance limits for sites with low exposure and narrower limits for sites with higher exposure. This method is different to the event rate based limits we have used in previous heuristics to measure {simaerep} performance. Funnel plots are discussed in greater detail by [Zink et al. 2018](https://doi.org/10.1177/2168479017738981)

# Prepare Data

## Load Portfolio Configurations
```{r real_config}

df_config <- readr::read_csv("ae_profile.csv")

df_config %>%
  head(25) %>%
  knitr::kable()
```

## Simulate Portfolio

And again we simulate artificial visit level data. Using parallel processing. At this stage we have simulated compliant test data of a realistic study portfolio.

```{r sim_portf}

df_portf <- sim_test_data_portfolio(df_config, parallel = TRUE, progress = TRUE)

df_portf %>%
  head(25) %>%
  knitr::kable()
```


# Apply {gsm}

## Example

```{r}
dfSUBJ <- df_portf %>%
  select(study_id, siteid = site_number, subjid = patnum, timeonstudy = visit) %>%
  summarise(timeonstudy = max(timeonstudy), .by = c(study_id, siteid, subjid)) %>%
  group_by(study_id) %>%
  nest()

dfAE <- df_portf %>%
  select(study_id, subjid = patnum, n_ae) %>%
  summarise(n_ae = max(n_ae), .by = c(study_id, subjid)) %>%
  filter(n_ae > 0) %>%
  mutate(n_ae = map(n_ae, ~ tibble(n = seq(1, .)), .progress = TRUE)) %>%
  unnest(n_ae) %>%
  select(- n) %>%
  group_by(study_id) %>%
  nest()
```

```{r}
dfInput <- gsm::AE_Map_Raw(list(dfSUBJ = dfSUBJ$data[[1]], dfAE = dfAE$data[[1]]))
dfInput

dfTransformed <- gsm::Transform_Rate(
  dfInput,
  strNumeratorCol = "Count",
  strDenominatorCol = "Exposure"
  )
dfTransformed

dfAnalyzed <- gsm::Analyze_NormalApprox(dfTransformed)
dfAnalyzed

dfFlagged <- gsm::Flag_NormalApprox(dfAnalyzed, vThreshold = c(-3, -2, 2, 3))
dfFlagged

dfSummary <- gsm::Summarize(dfFlagged)
dfSummary

dfBounds <- gsm::Analyze_NormalApprox_PredictBounds(dfTransformed, vThreshold = c(-3, -2, 2, 3))
dfBounds

chart <- gsm::Visualize_Scatter(dfFlagged, dfBounds)
chart
```

## Simulate UR

```{r}
sim_gsm_ur <- function(site, ur_rate, dfTransformed) {
  dfTransformed <- dfTransformed %>%
    mutate(
      Numerator = ifelse(GroupID == site, Numerator * (1 - ur_rate), Numerator),
      Metric = Numerator / Denominator
    )
  
  gsm::Analyze_NormalApprox(dfTransformed) %>%
    filter(GroupID == site) %>%
    pull(Score)
}

sim_gsm_ur("0001", ur_rate = 0.75, dfTransformed)
```


```{r}
df_sim <- dfSUBJ %>%
  inner_join(dfAE, by = "study_id") %>%
  ungroup() %>%
  mutate(
    trans = map2(data.x, data.y, ~ gsm::AE_Map_Raw(list(dfSUBJ = .x, dfAE = .y))),
    trans = map(trans, ~ gsm::Transform_Rate(., strNumeratorCol = "Count", strDenominatorCol = "Exposure")),
    sites = map(data.x, ~ distinct(., siteid))
  ) %>%
  select(- starts_with("data.")) %>%
  unnest(sites) %>%
  mutate(ur = list(tibble(ur_rate = c(0, 0.1, 0.25, 0.5, 0.75, 1)))) %>%
  unnest(ur) %>%
  mutate(
    score = pmap_dbl(list(siteid, ur_rate, trans), sim_gsm_ur, .progress = TRUE)
  )
  

df_sim
```

# UR {simaerep}

```{r sim_ur_scen_kable}
df_sim_simaerep <- sim_ur_scenarios(
   df_portf,
   extra_ur_sites = 0,
   ur_rate = c(0, 0.1, 0.25, 0.5, 0.75, 1),
   parallel = TRUE,
   poisson = TRUE,
   prob_lower = TRUE,
   progress = TRUE
)
```

# Evaluate

## Combine Results

```{r}

df_sim_thresh2 <- df_sim %>%
  mutate(
    is_ur = score <= -2,
    type = "{gsm} - thresh: -2",
    site_number = siteid
  ) %>%
  select(type, study_id, site_number, ur_rate, is_ur, score)


df_sim_simaerep_threshp95 <- df_sim_simaerep %>%
  mutate(
    is_ur = prob_low_prob_ur >= 0.95,
    type = "{simaerep} - thresh: 0.95"
  ) %>%
  select(type, study_id, site_number, ur_rate, is_ur, score = prob_low_prob_ur)


df_eval <- bind_rows(
  df_sim_thresh2,
  df_sim_simaerep_threshp95,
)
```


## Aggregate

```{r}
get_prop_test_ci95 <- function(..., ix) {
  
  stopifnot(ix %in% c(1, 2))
  
  tryCatch(
    prop.test(...)$conf.int[ix],
    error = function(cnd) c(NA, NA)[ix]
  )
}

aggr_results <- function(df_eval) {

df_perf <- df_eval %>%
  summarise(
    n = n(),
    .by = c(type, ur_rate, is_ur)
  ) %>%
  pivot_wider(
    names_from = is_ur,
    values_from = n,
    names_prefix = "is_ur_",
    values_fill = 0
  ) %>%
  mutate(
    n_sites = is_ur_TRUE + is_ur_FALSE + is_ur_NA,
    ratio = is_ur_TRUE / n_sites,
    ratio_type = ifelse(ur_rate == 0, "fpr", "tpr"),
    ci95_low = map2_dbl(is_ur_TRUE, n_sites, ~ get_prop_test_ci95(.x, .y, ix = 1)),
    ci95_high = map2_dbl(is_ur_TRUE, n_sites, ~ get_prop_test_ci95(.x, .y, ix = 2))
  )
}

df_perf <- aggr_results(df_eval)
```

## Table

```{r}
df_perf %>%
  knitr::kable(digits = 3)

```



## Plot Performance Metrics

- {gsm} has better performance than {simaerep}

```{r, plot, fig.width=6, fig.height = 10}

plot_perf <- function(df_perf) {

  df_perf %>%
    mutate(ur_rate = paste0("under-reporting rate: ",  ur_rate, " - ", ratio_type),
           ur_rate = ifelse(str_detect(ur_rate, "fpr"), "fpr", ur_rate)) %>%
    group_by(ur_rate) %>%
    ggplot(aes(type, ratio)) +
      geom_errorbar(aes(ymin = ci95_low, ymax = ci95_high), color = "violetred2") +
      facet_wrap(~ ur_rate, ncol = 1) +
      coord_flip() +
      labs(
        x = "",
        y = "CI95 Performance Ratio", 
        title = "{simaerep} vs {gsm} Performance",
        subtitle = "Only one under-reporting site per study."
      )
}

plot_perf(df_perf)
```


## Analysis

- {gsm} correctly flags more sites in studies with <= 0.25 AE per Visit with higher number of patients per site
- {simarep} correctly flags more sites in studies with 0.25 - 0.5 AE per visit with smaller number of patients per site
- false positive flagging for {simaerep} is similarly distributed as true positive flagging
- false positive flagging for {gsm} evenly covers the whole distribution range

```{r}
df_qu <- df_portf %>%
  summarize(n_pat = n_distinct(patnum), .by = c("study_id", "site_number", "ae_per_visit_mean")) %>%
  left_join(
    df_sim %>%
      select(study_id, site_number = siteid, ur_rate, score)
  ) %>%
  left_join(
    df_sim_simaerep %>%
      select(study_id, site_number, ur_rate, prob_low_prob_ur)
  ) %>%
  mutate(
    flag = case_when(
      score <= -2 & prob_low_prob_ur >= 0.95 ~ "double",
      score <= -2 ~ "gsm only",
      prob_low_prob_ur >= 0.95 ~ "simaerep only",
      TRUE ~ "miss"),
    flag = fct_relevel(flag,"simaerep only", "gsm only", "double", "miss")
  )

plot_flags <- function(df, ur_rate) {
  
  p <- df %>%
    filter(ur_rate == ur_rate) %>%
    ggplot(aes(log(n_pat), ae_per_visit_mean, colour = flag)) +
    geom_point(alpha = 0.25) +
    theme(legend.position = "bottom", aspect.ratio = 1) +
    labs(x = "Log Number of Patiends per Site", y = "AE per Visit Study Average", title = paste("Under Rporting Rate:", ur_rate))
  
  ggExtra::ggMarginal(p, groupColour = TRUE, groupFill = TRUE)
  
}

p_p5 <- plot_flags(df_qu, 0.5)

p_p75 <- plot_flags(df_qu, 0.75)
p_1 <- plot_flags(df_qu, 1)

p_fpr <-  df_qu %>%
  filter(ur_rate == 0, flag != "miss") %>%
  ggplot(aes(log(n_pat), ae_per_visit_mean, colour = flag)) +
  geom_point(alpha = 0.25) +
    theme(legend.position = "bottom", aspect.ratio = 1) +
    labs(x = "Log Number of Patiends per Site", y = "AE per Visit Study Average", title = paste("Fals Positive Flags"))

p_fpr <- ggExtra::ggMarginal(p_fpr, groupColour = TRUE, groupFill = TRUE)

gridExtra::grid.arrange(p_p5, p_p75, p_1, p_fpr, ncol = 2)
```

## Combine both approaches

- clear performance increase only for studies with > 1 ae per visit rate, which is very rare.

```{r}

cut_off <- 1

df_comb1 <- df_qu %>%
  mutate(
    type = "comb. ae/visit thresh: 1",
    is_ur = (ae_per_visit_mean <= 1 & score <= -2) | ae_per_visit_mean > 1 & prob_low_prob_ur >= 0.95
  ) %>%
  select(type, study_id, site_number, ur_rate, is_ur)

df_combp25 <- df_qu %>%
  mutate(
    type = "comb. ae/visit thresh: 0.25",
    is_ur = (ae_per_visit_mean <= 0.25 & score <= -2) | ae_per_visit_mean > 0.25 & prob_low_prob_ur >= 0.95
  ) %>%
  select(type, study_id, site_number, ur_rate, is_ur)

df_combp5 <- df_qu %>%
  mutate(
    type = "comb. ae/visit thresh: 0.5",
    is_ur = (ae_per_visit_mean <= 0.5 & score <= -2) | ae_per_visit_mean > 0.5 & prob_low_prob_ur >= 0.95
  ) %>%
  select(type, study_id, site_number, ur_rate, is_ur)


df_comb <- bind_rows(
  df_eval,
  df_comb1,
  df_combp25,
  df_combp5
)

df_comb_perf <- aggr_results(df_comb)

plot_perf(df_comb_perf) +
  labs(title = "Combined Performance with Differect ")
```


# Conclusion

- Across the portfolio {gsm} flagging performs better than {simaerep} flagging.
- A robust population of studies for which {simaerep} flagging is preferable cannot be identified.

The simulation assumes a constant AE rate during the the study. This is not true specifically during follow-up periods. By design {simaerep} should be able to handle those cases better than {gsm}. A more complex simulation is needed in order to proof this hypothesis.

[Zink et al. 2018](https://doi.org/10.1177/2168479017738981) addressees the limitation of funnel plots for changing event rates in different phases of the study and suggests to stratify the data by relative time windows and present the results as a heatmap.



```{r close}
plan(sequential)
```
