---
title: "Inframe Simulation Using Table Operations"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    collapse: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, message = FALSE)
```

# Load
```{r load}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(simaerep))
```

# Introduction

With the latest `0.6.0` release we have added an alternative version of the {simaerep} algorithm that was coded using solely `dbplyr` compatible table operations.

- expand the patients for each site `r` times
- join each patient with a random eligible patient from same study
- for each replicate calculate event per visit rate per site
- calculate the ratio of having a lower event per visit rate than actually observed

This comes with the following advantages and disadvantages:

- Patients are individually matched with patients that have reached the same visit in the study. No need to pick visit_med75 as an evaluation point.
- `dbplyr` compatibility means that code execution can be done in a database back-end as opposed to in-memory.
- Matching patients individually is more costly, this increases in-memory computation time
- Limited patient sample pool for patients that have more visits than other patients in study.

# Sample Data

```{r}
set.seed(1)

df_visit <- sim_test_data_study(
  n_pat = 1000, # number of patients in study
  n_sites = 100, # number of sites in study
  frac_site_with_ur = 0.05, # fraction of sites under-reporting
  ur_rate = 0.4, # rate of under-reporting
  ae_per_visit_mean = 0.5 # mean AE per patient visit
)

df_visit$study_id <- "A"
```

# Patient-Level Matching

Here we use the standard version of the algorithm.

```{r}
aerep_trad <- simaerep(df_visit)
plot(aerep_trad)
```

To use the patient-level matching algorithm we set `inframe=TRUE` and `visit_med75=FALSE`.

The original algorithm uses fixed seeds before sampling while the inframe method does not. In order to obtain consistent results we need to manually set a seed.

```{r}
set.seed(1)

aerep_inframe <- simaerep(
  df_visit,
  inframe = TRUE,
  visit_med75 = FALSE
)
```

The plot shows that for all sites 10/10 patients were used and none were excluded. We also observe that
the site average has become more noisy as less patients are used to calculate the averages for the higher
visit numbers.

```{r}
plot(aerep_inframe)
```

The inframe method includes this noisier data but does not compare average event counts but event per visit rates.
We can find `events_per_visit_site` and `events_per_visit_study` in `df_eval`. The latter is the average event rate
obtained in the simulation in which each patient has been resampled according to its maximum visit.

```{r}
aerep_inframe$df_eval
```

We can also force the inframe method to use the visit_med75 this will prefilter `df_visit`, which adds an extra step
and decreases performance. 

```{r}

set.seed(1)

aerep_inframe_visit_med75 <- simaerep(
  df_visit,
  inframe = TRUE,
  visit_med75 = TRUE
)

plot(aerep_inframe_visit_med75)
```


# DB

We can demonstrate the database-backend compatibility by using a connection to a in memory `duckdb` database. In order to set the number of replications we need to create a new table in our back-end that has one column with as many rows as the desired replications.

A lazy reference to this table can then be passed to the `r` parameter.

```{r}
con <- DBI::dbConnect(duckdb::duckdb(), dbdir = ":memory:")
df_r <- tibble(rep = seq(1, 1000))

dplyr::copy_to(con, df_visit, "visit")
dplyr::copy_to(con, df_r, "r")

tbl_visit <- tbl(con, "visit")
tbl_r <- tbl(con, "r")


aerep <- simaerep(tbl_visit, r = tbl_r, visit_med75 = FALSE)
```

When inspecting `df_eval` we see that it is still a lazy table object.

```{r}
aerep$df_eval
```

We can convert it to sql code. The `cte` option makes the sql code more readable.

```{r}
sql_eval <- dbplyr::sql_render(aerep$df_eval, sql_options = dbplyr::sql_options(cte = TRUE))
stringr::str_trunc(sql_eval, 500)
```

We can take that code and wrap it in a `CREATE TABLE` statement

```{r}
sql_create <- glue::glue("CREATE TABLE eval AS ({sql_eval})")
DBI::dbExecute(con, sql_create)
```

Retrieve the new table from the database.

```{r}
tbl_eval <- tbl(con, "eval")
tbl_eval
```

We plot the results from the {simaerep} object.

```{r}
plot(aerep)
```

Or more efficiently by using `plot_study()` when we have already written the `simaerep` results into the database.
Here we avoid that the results are being recalculated just for the sake of creating a plot. However this requires
that we save `df_site` to the database as well.

```{r}
sql_site <- dbplyr::sql_render(aerep$df_site)
DBI::dbExecute(con, glue::glue("CREATE TABLE site AS ({sql_site})"))
tbl_site <- tbl(con, "site")

plot_study(tbl_visit, tbl_site, tbl_eval, study = "A")
```


```{r}
DBI::dbDisconnect(con)
```



# In Memory Calculation Times

Here we perform some examplary tests to illustrate the increase in-memory calculation time of the inframe calculation method.

This is the calculation time for the default settings
```{r}
system.time({simaerep(df_visit, inframe = FALSE, visit_med75 = TRUE, under_only = TRUE, progress = FALSE)})
```


inframe calculation time is higher

```{r}
system.time({simaerep(df_visit, inframe = TRUE, visit_med75 = FALSE, under_only = TRUE)})
```


