---
title: "Over-Reporting Probability"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    collapse: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, fig.width = 10)
```

# Load
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(simaerep))
suppressPackageStartupMessages(library(ggExtra))
```

# Introduction

{siamerep} was originally created to detect under-reporting of AEs and therefore no over-reporting probability was calculated. Nevertheless {simaerep} can theoretically be used to simulate all kinds of subject-based clinical events, for some such as issues over-reporting can represent a quality issue.  With the recent release `0.5.0` we have added the option to calculate an over-reporting probability score.

# Data Set

We simulate a standard data set with a high number of sites, patients, visits and events to ensure that most of our dimensions will be normally distributed. We do not add any over- or under-reporting sites at this point.

```{r}
df_visit <- sim_test_data_study(
 n_pat = 10000,
 n_sites = 1000,
 frac_site_with_ur = 0,
 max_visit_mean = 100,
 max_visit_sd = 1,
 ae_per_visit_mean = 5
)

df_visit$study_id <- "A"
```

# Run {simaerep}

in order to add the over-reporting probability score we need to set the parameter `under_only = FALSE`.

```{r}
system.time(
  aerep_def <- simaerep(
    df_visit,
    param_sim_sites = list(under_only = TRUE),
    param_eval_sites = list(under_only = TRUE)
  )
)
```

The original setting skips the simulation for all sites that do have more AEs than the study average.

```{r}
system.time(
  aerep_ovr <- simaerep(
    df_visit,
    param_sim_sites = list(under_only = FALSE),
    param_eval_sites = list(under_only = FALSE)
  )
)
```

The new parameter calculates the probability of a site getting `a lower or equal` average AE count for the site visit_med75 for every site, regardless of how its initial value compares to the study average. The calculation only takes a few seconds longer than the default setting.

In the evaluation data frame we have three more columns available now.

```{r}
setdiff(colnames(aerep_ovr$df_eval), colnames(aerep_def$df_eval))
```

# Analyze

## Probability getting a lower or equal AE count


```{r}

cols <- c("study_id", "site_number", "mean_ae_site_med75", "mean_ae_study_med75", "prob_low")

p <- bind_rows(
  select(
      aerep_ovr$df_eval,
      all_of(cols)
    ) %>%
    mutate(type = "with over-reporting"),
    select(
      aerep_def$df_eval,
      all_of(cols)
    ) %>%
    mutate(type = "default")
  ) %>%
  ggplot(aes(x = mean_ae_site_med75 - mean_ae_study_med75, y = prob_low, color = type)) +
    geom_point(alpha = 0.5) +
    theme(legend.position = "bottom") +
    scale_color_manual(values = c("gold", "blue")) +
    labs(y = "Probability of getting a lower or equal mean site AE in a 1000x simulation")

ggExtra::ggMarginal(p, groupColour = TRUE, type = "density")
```


We can see that we have a gap for the default setting in the generated probabilities. The values filling the gap can be interpreted as the probability of having a `higher` site average than originally observed.

We need to be aware that values in the simulation `equal` to the site mean will count towards the under-reporting probability. This can potentially lead to artifacts with low AE counts.

## Over-Reporting

We can add the over-reporting probability as (1- under-reporting probability)

```{r}

cols <- c("study_id", "site_number", "mean_ae_site_med75", "mean_ae_study_med75")


p <- bind_rows(
  select(
      aerep_ovr$df_eval,
      all_of(cols),
      value = "prob_low"
    ) %>%
    mutate(type = "new under-reporting"),
  select(
      aerep_ovr$df_eval,
      all_of(cols),
      value = "prob_high"
    ) %>%
    mutate(type = "new over-reporting"),
    select(
      aerep_def$df_eval,
      all_of(cols),
      value = "prob_low"
    ) %>%
    mutate(type = "default under-reporting")
  ) %>%
  ggplot(aes(x = mean_ae_site_med75 - mean_ae_study_med75, y = value, color = type)) +
    geom_point(alpha = 0.25) +
    theme(legend.position = "bottom") +
    scale_color_manual(values = c("gold", "purple", "blue")) +
    labs(y = "Probability of getting a lower or equal mean site AE in a 1000x simulation")

ggExtra::ggMarginal(p, groupColour = TRUE, type = "density")
```

## Multiplicity Correction

The multiplicity correction dampens the signal, avoiding false positives that are the result of chance.

```{r}

cols <- c("study_id", "site_number", "mean_ae_site_med75", "mean_ae_study_med75")


p <- bind_rows(
  select(
      aerep_ovr$df_eval,
      all_of(cols),
      value = "prob_low_prob_ur"
    ) %>%
    mutate(type = "new under-reporting"),
  select(
      aerep_ovr$df_eval,
      all_of(cols),
      value = "prob_high_prob_or"
    ) %>%
    mutate(type = "new over-reporting"),
    select(
      aerep_def$df_eval,
      all_of(cols),
      value = "prob_low_prob_ur"
    ) %>%
    mutate(type = "default under-reporting")
  ) %>%
  ggplot(aes(x = mean_ae_site_med75 - mean_ae_study_med75, y = value, color = type)) +
    geom_point(alpha = 0.25) +
    theme(legend.position = "bottom") +
    scale_color_manual(values = c("gold", "purple", "blue")) +
    labs(y = "Probability of getting a lower or equal mean site AE in a 1000x simulation")

ggExtra::ggMarginal(p, groupColour = TRUE, type = "density")
```

# Simulating Over-Reporting

We can simulate under-reporting by supplying a negative ratio for `ur_rate`

```{r}
df_visit <- sim_test_data_study(
 frac_site_with_ur = 0.05,
 ur_rate = - 0.5,
)

df_visit$study_id <- "A"


distinct(df_visit, site_number, is_ur, ae_per_visit_mean)
```


```{r}
aerep <- simaerep(
    df_visit,
    param_sim_sites = list(under_only = FALSE),
    param_eval_sites = list(under_only = FALSE)
  )
```

```{r}
aerep$df_eval %>%
  select(site_number, mean_ae_site_med75, mean_ae_study_med75, prob_low_prob_ur, prob_high_prob_or)
```

We can plot over-reporting by changing setting `prob_col = "prob_high_prob_or"`.

```{r}
plot(aerep, prob_col = "prob_high_prob_or")

plot(aerep, prob_col = "prob_low_prob_ur")

```

