---
title: "Adjusted Evaluation Point visit_med75"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    collapse: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
```

# Load
```{r}
suppressPackageStartupMessages( library(tidyverse) )
suppressPackageStartupMessages( library(knitr) )
suppressPackageStartupMessages( library(simaerep) )
```

# Introduction

Previously we used an unadjusted visit_med75. Simply taking the median of all maximum patient visits at a site and multiplying by 0.75. If the number of patients at a site is reasonably large the smallest maximum visit of the patients making the cut-off will be close or equal to the cut-off point.

```{r}

set.seed(1)

df_visit <- sim_test_data_study(
  n_pat = 120,
  n_sites = 6,
  frac_site_with_ur = 0.4,
  ur_rate = 0.6)

df_visit$study_id <- "A"
df_site <- site_aggr(df_visit, method = "med75_ad")

plot_visit_med75(df_visit, df_site, study_id_str = "A", n_site = 6)
```

# Only few Patients per Site

If we only have a few patients per site chances will be greater that the smallest maximum visit of those patients that make the cut-off is much greater than the cut-off point and we could thus inlclude more visits into our analysis. For those cases we adjust the evaluation point to the smallest maximum visit of all patients that make the initial cut.

```{r}
set.seed(1)

df_visit <- sim_test_data_study(
  n_pat = 24,
  n_sites = 6,
  frac_site_with_ur = 0.4,
  ur_rate = 0.6)

df_visit$study_id <- "A"
df_site <- site_aggr(df_visit, method = "med75_ad")

plot_visit_med75(df_visit, df_site, study_id_str = "A", n_site = 6)

```

# Early Starters

It can happen that we have patients that are recruited very early in the study which are also located at a few early starting sites.  

```{r}
  # create three small sites with early and late starting patients
  df_visit1 <- sim_test_data_study(
    n_pat = 18,
    n_sites = 3,
    frac_site_with_ur = 0.4,
    ur_rate = 0.2,
    max_visit_sd = 2,
    max_visit_mean = 20
  )

  df_visit2 <- sim_test_data_study(
    n_pat = 9,
    n_sites = 3,
    frac_site_with_ur = 0.4,
    ur_rate = 0.2,
    max_visit_sd = 2,
    max_visit_mean = 5
  )


  df_visit1$patnum <- paste0("A", df_visit1$patnum)
  df_visit2$patnum <- paste0("B", df_visit2$patnum)

  df_visit <- bind_rows(df_visit1, df_visit2)

  # create three larger sites with only late starting patients
  df_visit3 <- sim_test_data_study(
    n_pat = 60,
    n_sites = 3,
    frac_site_with_ur = 0.4,
    ur_rate = 0.2,
    max_visit_sd = 2,
    max_visit_mean = 10
  )

  df_visit3$site_number <- paste0("C", df_visit3$site_number)

  df_visit <- bind_rows(df_visit, df_visit3)

  df_visit$study_id <- "A"

  df_site <- site_aggr(df_visit)

  study_qup8_max_visit <- df_visit %>%
    group_by(patnum) %>%
    summarise(max_visit = max(visit)) %>%
    pull(max_visit) %>%
    quantile(0.8) %>%
    round()

  testthat::expect_true(max(df_site$visit_med75) <= study_qup8_max_visit)

  # nolint start
  # plot_visit_med75(df_visit, df_site, study_id_str = "A", n_site = 6)
  # nolint end

```

