---
title: "Aggregate AEs by Days or Visit?"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    collapse: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, fig.width = 10)
```

# Load
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(simaerep))
```

# Introduction

We generally aggregate AEs by visit. Each patient follows the same visit schedule and a specified number of days passes between each consecutive visit. As visits are a scheduled contact point between the patient and physicians the day of the visit usually also the day when most AEs get reported. Alternatively we can also choose to use `simaerep` to a


# Load Data

We load some a public clinical trial data set which only contains data of the control arm [see SAS files as a Data Source Article](https://openpharma.github.io/simaerep/articles/sas_files.html)

```{r}
df_ae <- haven::read_sas('adae.sas7bdat') %>%
  select(STUDYID, SUBJID, SITEID, AESTDY) 

df_vs <- haven::read_sas('advs.sas7bdat') %>%
  select(STUDYID, SUBJID, SITEID, ADY) 

df_ae <- df_ae %>%
  rename(DY = AESTDY) %>%
  mutate(EVENT = "AE")

df_vs <- df_vs %>%
  rename(DY = ADY) %>%
  mutate(EVENT = "VS") %>%
  # we ignore visits that have no date
  filter(! is.na(DY)) %>%
  # we are not interested in same day visits
  distinct()

df_aevs <- bind_rows(df_ae, df_vs) %>%
  # NA's get sorted towards the end thus AEs with no date get sorted towards last visit
  arrange(STUDYID, SITEID, SUBJID, DY) %>%
  group_by(STUDYID, SITEID, SUBJID) %>%
  mutate(AE_NO = cumsum(ifelse(EVENT == "AE", 1, 0)),
         VS_NO = cumsum(ifelse(EVENT == "VS", 1, 0))) %>%
  # we remove patients with 0 visits
  filter(max(VS_NO) > 0) %>%
  # AE's before fist visit should register to visit 1 not zero
  mutate(VS_NO = ifelse(VS_NO == 0, 1, VS_NO))


df_aevs_aggr <- df_aevs %>%
  group_by(STUDYID, SITEID, SUBJID, VS_NO) %>%
  summarise(MIN_AE_NO = min(AE_NO),
            MAX_AE_NO = max(AE_NO),
            .groups = "drop") %>%
  group_by(STUDYID, SITEID, SUBJID) %>%
  mutate(MAX_VS_PAT = max(VS_NO)) %>%
  ungroup() %>%
  # assign AEs that occur after last visit to last AE
  mutate(
    CUM_AE = ifelse(
      VS_NO == MAX_VS_PAT,
      MAX_AE_NO,
      MIN_AE_NO)
    )

df_visit <- df_aevs_aggr %>%
  rename(
    study_id = "STUDYID",
    site_number = "SITEID",
    patnum = "SUBJID",
    n_ae = "CUM_AE",
    visit = "VS_NO"
  )  %>%
  select(study_id, site_number, patnum, n_ae, visit)

```


# Aggregate on Days

For aggregating on days we need to align the reference timelines of the single patients.

```{r}
df_vs_min_max <- df_vs %>%
  group_by(STUDYID, SUBJID, SITEID) %>%
  summarise(min_DY = min(DY, na.rm = TRUE),
            max_DY = max(DY, na.rm = TRUE),
            .groups = "drop")

df_vs_min_max$min_DY[1:25]
df_vs_min_max$max_DY[1:25]

```

The day of the first visit is different for each patient and they start at negative values.
First we correct all values to be positive and then normalize the AE date values to the date value of the first visit of each patient

```{r}
corr_factor <- abs(min(df_vs_min_max$min_DY))

df_days <- df_ae %>%
  # include patients with vs but no AE
  right_join(df_vs_min_max, by = c("STUDYID", "SUBJID", "SITEID")) %>%
  # replace DY NULL with max patient DY 
  group_by(STUDYID, SUBJID, SITEID) %>%
  mutate(DY = ifelse(is.na(DY) & ! is.na(EVENT), max(DY, na.rm = TRUE), DY)) %>%
  # replace DY for patients with 0 AE with day of maximum visit
  mutate(DY = ifelse(is.na(DY) & is.na(EVENT), max_DY, DY)) %>%
  # correct timelines
  mutate(DY = DY + corr_factor,
         min_DY = min_DY + corr_factor,
         DY_corr = DY + min_DY) %>%
  group_by(STUDYID, SITEID, SUBJID) %>%
  arrange(STUDYID, SITEID, SUBJID, DY_corr) %>%
  mutate(n_ae = row_number()) %>%
  ungroup() %>%
  # set AE count to 0 for patients with no AEs
  mutate(n_ae = ifelse(is.na(EVENT), 0 , n_ae)) %>%
  rename(
    study_id = STUDYID,
    site_number = SITEID,
    patnum = SUBJID,
    visit = DY_corr
  ) %>%
  group_by(study_id, site_number, patnum, visit) %>%
  summarise(n_ae = max(n_ae), .groups = "drop")
```

check if we get the same transformation as for the visit aggregations

```{r}

stopifnot(n_distinct(df_days$site_number) == n_distinct(df_visit$site_number))
stopifnot(n_distinct(df_days$patnum) == n_distinct(df_visit$patnum))

pat0_days <- df_days %>%
  group_by(study_id, site_number, patnum) %>%
  filter(max(n_ae) == 0) %>%
  pull(patnum) %>%
  unique() %>%
  sort()

pat0_vs <- df_visit %>%
  group_by(study_id, site_number, patnum) %>%
  filter(max(n_ae) == 0) %>%
  pull(patnum) %>%
  unique() %>%
  sort()

stopifnot(all(pat0_days == pat0_vs))
```


```{r}
df_days 
```

We do have gaps in between the days leading to implicitly missing values. `simaerep` will correct this automatically and throw a warning.

```{r}
df_site <- site_aggr(df_visit = df_days)
```

to silence the warning we can use `check df_visit()` which is also called internally by all other functions accepting `df_visit` as an argument.

```{r}
df_days <- simaerep:::check_df_visit(df_days)

df_days
```

Then we proceed as usual.

```{r}
df_sim_sites <- sim_sites(df_site, df_visit = df_days)

df_eval_days <- eval_sites(df_sim_sites)

simaerep::plot_study(df_visit = df_days, df_site = df_site, df_eval = df_eval_days, study = unique(df_days$study_id))
```


# Aggregate on Visits

How do the results compare to aggregating on visits?

```{r}

df_site <- site_aggr(df_visit) 

df_sim_sites <- sim_sites(df_site, df_visit)

df_eval_vs <- eval_sites(df_sim_sites)

simaerep::plot_study(df_visit, df_site, df_eval_vs, study = unique(df_visit$study_id))
```

# Compare

We observe a difference in the results. Which is largely attributable in the difference in cut-off visit_med75 points that influences the set of patients included. In any case we observe a high rank correlation with a low p-value of all results greater 0.

As the inclusion/exclusion of patients in the analysis of a site in an ongoing trial can shift results, we recommend to aggregate on actually occurred visits because then all included patients had an equal amount of opportunities to report AEs.

```{r}
df_comp <- df_eval_days %>%
  select(
    site_number,
    prob_low_prob_ur_days = prob_low_prob_ur,
    n_pat_with_med75_days = n_pat_with_med75
  )  %>%
  left_join(
    select(
      df_eval_vs,
      site_number,
      prob_low_prob_ur_vs = prob_low_prob_ur,
      n_pat_with_med75_vs = n_pat_with_med75
      ),
    by = "site_number"
  ) %>%
  filter(prob_low_prob_ur_days > 0 | prob_low_prob_ur_vs > 0) %>%
  select(site_number, starts_with("prob"), starts_with("n_pat")) %>%
  arrange(desc(prob_low_prob_ur_vs))

df_comp %>%
  DT::datatable()

cor.test(
  df_comp$prob_low_prob_ur_vs,
  df_comp$prob_low_prob_ur_days,
  method = "spearman"
)
```

